\sysfull

### 2018-08-07 paper/asplos19/abstract.tex
revisions of abstract and intro (Sinan's comments applied)

Energy-neutral Internet of Things requires freeing embedded devices from batteries and powering them from ambient energy. Ambient energy is, however, unpredictable and can only power a device intermittently. Therefore, the paradigm of intermittent execution is to save the program state into non-volatile memory frequently to preserve the execution progress. In task-based intermittent programming the state is saved at task transition. Tasks are fixed at compile-time and agnostic to energy conditions. Thus, the state may be saved either more often than necessary or not often enough for the program to progress and terminate. To address this challenges, we propose \sys, an adaptive and efficient task-based execution model. \sys progresses on a multi-task scale when energy permits and preserves the computation progress on a sub-task scale if necessary. \sys's specialized memory virtualization mechanism ensures that power failures do not leave the program state in non-volatile memory inconsistent. Our evaluation on a real energy-harvesting platform not only shows that \sys reduces run time by up to 54\%, and by 26\% on average, as compared to a state-of-the-art system, but also it is able to \emph{progress} where static systems \emph{fail}.

>>>>

Energy-neutral Internet of Things requires freeing embedded devices from batteries and powering them from ambient energy. Ambient energy is, however, unpredictable and can only power a device intermittently. Therefore, the paradigm of intermittent execution is to save the program state into non-volatile memory frequently to preserve the execution progress. In task-based intermittent programming the state is saved at task transition. Tasks are fixed at compile-time and agnostic to energy conditions. Thus, the state may be saved either more often than necessary or not often enough for the program to progress and terminate. To address this challenges, we propose \sys, an adaptive and efficient task-based execution model. \sys progresses on a multi-task scale when energy permits and preserves the computation progress on a sub-task scale if necessary. \sys's specialized memory virtualization mechanism ensures that power failures do not leave the program state in non-volatile memory inconsistent. Our evaluation on a real energy-harvesting platform not only shows that \sys reduces run time by up to 54\% as compared to a state-of-the-art system, but also it is able to \emph{progress} where static systems fail.

### 2018-08-07 paper/asplos19/intro.tex
revisions of abstract and intro (Sinan's comments applied)

Upon power failures, a device loses the volatile state in its registers, stack, SRAM, and retains the state of any non-volatile memory, such as FRAM. While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave non-volatile state incorrect, partially updated. These inconsistencies cause intermittent execution to deviate from continuously-powered behavior, leading to an unrecoverable application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for intermittently-powered devices: (i) \emph{software-based programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{hardware-based architectural support}~\cite{hicks_isca_2017,idetic,nvp}. Complex hardware architectural changes are expensive to design, verify, and manufacture. New hardware architectures are also inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software approaches are simpler and applicable to existing devices today. Therefore, this work focuses on software approaches. In particular we address the key limitation of task-based programming and execution model, that is the {\em inflexibility} and \emph{energy-unawareness} of statically decomposing a program into tasks.
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/intro-figure-v5.pdf}
    \caption{Task coalescing reduces commit overhead (left), while task splitting enables termination of large tasks (right).
    S: static legacy task-based runtime, C: \sys.}
    % \caption{Task coalescing reduces time and energy overhead by performing fewer commits (denoted as \texttt{C}) than the individual tasks (denoted as \texttt{Tx}) commits (left). On the other hand, task splitting reduces wasted computation and enables termination of big tasks, requiring more energy than the energy buffer can hold (right).}
    \label{fig:coalesce}
\end{figure}

@@

\paragraph{Challenges and Contributions.}
We introduce \sys: a new task-based system that employs \emph{adaptive task size execution by task coalescing and splitting}. By means of this novel technique, small tasks can be executed efficiently by trimming unnecessary overhead \emph{dynamically}, meanwhile avoiding the risk of non-termination. \sys accepts any static decomposition and it coalesces (groups) tasks or splits them (see again Fig.~\ref{fig:coalesce}) based on the estimated energy availability without demanding any hardware support. To the best of our knowledge, \sys is the only system that guarantees forward progress for any statically-defined task  set or energy storage.

@@

\item[C3] \emph{Ensuring Task Termination:} a static task decomposition model assumes that each task can execute to completion. If the hardware energy buffer provides inadequate energy to execute each task to completion, a program will not terminate~\cite{cleancut_2018}. This leads to a third challenge: how to enable the dynamic execution model to progress on a sub-task level? To avoid non-termination under adverse energy conditions, \sys uses a novel timer-based {\em partial task commit} mechanism, inspired by~\cite{ratchet}. Partial execution avoids non-termination by committing the intermediate state of a long-running task that has repeatedly failed and restarted.

@@

To asses the benefits of \sys over existing task-based systems, we implemented and tested six benchmarks on a real energy-harvesting platform. Our evaluation not only shows that \sys reduces run time by up to 54\%, and by 26\% on average, but also is able to continue the execution where static task-based systems fail.

The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.
Section~\ref{sec:systemdescription} illustrates an overview of \sys. Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism. The dynamic page privatization is explained in Section~\ref{sec:memory_virtulaization}. Implementation details of \sys are given in Section~\ref{sec:implementation}. Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's evaluation methodology and results. Section~\ref{sec:related_work} positions \sys in the context of related work and Section~\ref{sec:conclusions} concludes and discusses future work.

>>>>

Upon power failures, a device loses the volatile state in its registers, stack, SRAM, and retains the state of any non-volatile memory, such as FRAM. While capturing periodic checkpoints~\cite{mementos,quickrecall} and sleep scheduling~\cite{dewdrop,hibernus,hibernusplusplus} help preserve execution progress, failures can leave non-volatile state incorrect, partially updated. These inconsistencies cause intermittent execution to deviate from continuously-powered behavior, leading to an unrecoverable application failure~\cite{dino,edb}. Prior work developed two main approaches to deal with data inconsistency for intermittently-powered devices: (i) \emph{software-based programming and execution models}~\cite{dino,ratchet,chain,alpaca} and (ii) \emph{hardware-based architectural support}~\cite{hicks_isca_2017,idetic,nvp}. Complex hardware architectural changes are expensive to design, verify, and manufacture. New hardware architectures are also inapplicable to existing systems~\cite{hicks_isca_2017,nvp}. Software approaches are simpler and applicable to existing devices today. Therefore, this work focuses on software approaches. In particular we address the key limitation of task-based programming and execution model, that is the {\em inflexibility} and \emph{energy-unawareness} of statically decomposing a program into tasks.
%
\begin{figure}
    \centering
    \includegraphics[width=\columnwidth]{figures/intro-figure-v5.pdf}
    \caption{Task coalescing reduces commit overhead (left), while task splitting enables termination of large tasks (right).
    S: static legacy task-based runtime, C: \sys (this work).}
    % \caption{Task coalescing reduces time and energy overhead by performing fewer commits (denoted as \texttt{C}) than the individual tasks (denoted as \texttt{Tx}) commits (left). On the other hand, task splitting reduces wasted computation and enables termination of big tasks, requiring more energy than the energy buffer can hold (right).}
    \label{fig:coalesce}
\end{figure}

@@

\paragraph{Challenges and Contributions.}
We introduce \sys: a new task-based system that employs \emph{adaptive task size execution by task coalescing and splitting}. By means of this novel technique, small tasks can be executed efficiently by trimming unnecessary overhead \emph{dynamically}, meanwhile avoiding the risk of non-termination. \sys accepts any static decomposition and it coalesces (groups) tasks or splits them (see again Fig.~\ref{fig:coalesce}) based on the estimated energy availability without demanding any hardware support. To the best of our knowledge, \sys is the only system that eliminates restructuring and re-compilation of applications considering device's energy buffer---enabling energy-storage independency for the applications, while keeping execution efficient.

@@

\item[C3] \emph{Ensuring Task Termination:} a static task decomposition model assumes that each task can execute to completion. If the hardware energy buffer provides inadequate energy to execute each task to completion, a program will not terminate~\cite{cleancut_2018}. This leads to a third challenge: how to enable the dynamic execution model to progress on a sub-task level? To avoid non-termination under adverse energy conditions, \sys uses a timer-based {\em partial task commit} mechanism. Partial execution avoids non-termination by committing the intermediate state of a long-running task that has repeatedly failed and restarted.

@@

To asses the benefits of \sys over existing task-based systems, we implemented and tested six benchmarks on a real energy-harvesting platform. Our evaluation shows that \sys reduces run time overhead by up to 54\% and solves task non-termination problem where existing static task-based systems fail.

The rest of this paper is organized as follows. Section~\ref{sec:background} provides background on intermittent computing.
Section~\ref{sec:systemdescription} illustrates an overview of \sys. Section~\ref{sec:task_adaptation} describes \sys's task adaptation mechanism. The page privatization is explained in Section~\ref{sec:memory_virtulaization}. Implementation details of \sys are given in Section~\ref{sec:implementation}. Sections~\ref{sec:methodology} and~\ref{sec:evaluation} describe \sys's evaluation methodology and results, respectively. Section~\ref{sec:related_work} positions \sys in the context of related work and Section~\ref{sec:conclusions} concludes our work.
